# Modeling {background-color='{{< brand color secondary >}}'}

```{r}
#| label: pkg-setup-05
#| warning: false
#| message: false
#| include: false
source("setup.R")
options(width = 77)
```

## Introduction

::: incremental

* Modeling investigates the relationship between a single dependent variable and one or more independent variables.
* **Logistic regression** is a specific case of the generalized linear model (GLM).
* Logistic regression is used to model binary outcomes, such as whether or not someone voted.

:::

## Introduction

Assumptions in logistic regression using survey data include:

* The outcome variable has two levels
* There is a linear relationship between the independent variables and the log odds (the equation for the logit function)
* The residuals are homoscedastic; that is, the error term is the same across all values of independent variables

## `svyglm`: syntax

```{.r code-line-numbers="|1|2,8|3|4|5|6|7"}
des_obj %>%
  svyglm(
    formula = outcomevar ~ x1 + x2 + x3,
    design = .,
    na.action = na.omit,
    df.resid = NULL,
    family = quasibinomial
  )
```

## `svyglm`: example

Model whether a person usually trusts the government based on who they voted for in 2024[^1]:

. . .

1. Create a binary trust variable by grouping ‘Always’ and ‘Most of the time’ into one category, and all remaining responses into the other.

```{r}
#| echo: true
anes_des_der <- anes_des %>%
  mutate(TrustGovernmentUsually = case_when(
    is.na(TrustGovernment) ~ NA,
    TRUE ~ TrustGovernment %in% c("1. Always", "2. Most of the time")
  ))
```

[^1]: Harris, Trump, or Other

## `svyglm`: example

2. Summarize the data:

```{r}
#| echo: true
#| code-line-numbers: "|3|4-10"
anes_des_summary <-
  anes_des_der %>%
  group_by(VotedPres2024_selection) %>%
  summarize(pct_trust = survey_mean(
    TrustGovernmentUsually,
    na.rm = TRUE,
    proportion = TRUE,
    vartype = "ci"),
    .groups = "drop"
    ) %>%
  filter(complete.cases(.))
```

## `svyglm`: example

3. Plot the summary of the data:

```{r}
#| dpi: 300
#| code-fold: true
anes_des_summary %>%
  ggplot(aes(
    x = VotedPres2024_selection, y = pct_trust,
    fill = VotedPres2024_selection
  )) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = pct_trust_low, ymax = pct_trust_upp),
    width = .2
  ) +
  scale_fill_manual(values = c("#0b3954", "#bfd7ea", "#8d6b94")) +
  xlab("Election choice (2020)") +
  ylab("Usually trust the government") +
  scale_y_continuous(labels = scales::percent) +
  guides(fill = "none") +
  theme_minimal()
```

## `svyglm`: example

4. Verify by fitting the model:

```{r}
#| output: slide
#| echo: true
#| code-line-numbers: "|2,6|3|4|5"
logistic_trust_vote <- anes_des_der %>%
  svyglm(
    design = .,
    formula = TrustGovernmentUsually ~ VotedPres2024_selection,
    family = quasibinomial
  )
```

## `svyglm`: example

4. Verify by fitting the model:

```{r}
#| echo: true
#| code-line-numbers: "|1,5|2,6|3,7,8"
library(broom)
library(prettyunits)
library(gt)

tidy(logistic_trust_vote) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

## `svyglm`: example

Result:

```{r}
tidy(logistic_trust_vote) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

1. Estimated coefficients (`estimate`)
2. Estimated standard errors of the coefficients (`std.error`)
3. p-value for each coefficient

## `svyglm`: example

Result:

```{r}
tidy(logistic_trust_vote) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

Compared with Harris voters, Trump voters show lower levels of trust in the government.

. . .

The coefficient of **-1.24** represents the increase in the log odds of usually trusting the government.

## `svyglm`: example

To talk about odds rather than log odds, we can exponentiate the coefficients:

```{r}
#| code-line-numbers: "|1"
#| echo: true
tidy(logistic_trust_vote, exponentiate = TRUE) %>%
  select(term, estimate) %>%
  gt() %>%
  fmt_number()
```

## `svyglm`: example

Result:

```{r}
#| code-line-numbers: "|1"
tidy(logistic_trust_vote, exponentiate = TRUE) %>%
  select(term, estimate) %>%
  gt() %>%
  fmt_number()
```

The odds that a Trump voter usually trusts the government are about **29%** of the odds for a Harris voter, who serves as the reference group.

## Interaction effects

Next, examine the interaction between early voting and level of education in the model.

1. Subset to 2024 voters and create an indicator for Harris voters.

```{r}
#| echo: true
anes_des_ind <- anes_des %>%
  filter(!is.na(VotedPres2024_selection)) %>%
  mutate(VoteHarris = case_when(
    VotedPres2024_selection == "1. Kamala Harris" ~ 1,
    TRUE ~ 0
  ))
```

## Interaction effects

2. Let’s look at the main effects of education and early voting behavior.

```{r}
#| echo: true
#| code-line_numbers: "|7"
log_harris_main <- anes_des_ind %>%
  mutate(
    EarlyVote2024 = fct_relevel(EarlyVote2024, "2. Have not voted", after = 0)
  ) %>%
  svyglm(
    design = .,
    formula = VoteHarris ~ EarlyVote2024 + EducationGroup,
    family = quasibinomial
  )
```

## Interaction effects {.smaller}

Results:

```{r}
#| echo: true
tidy(log_harris_main) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

## Interaction effects {.smaller}

Results:

```{r}
#| echo: false
tidy(log_harris_main) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

This indicates that people with Bachelor's and graduate degrees were more likely to vote for Harris in the 2024 election compared to people with less than a high school diplomas (reference level).

. . .

Early voting behavior was not significant.

## Interaction effects

To find the interaction between education level and early voting behavior:

```{r}
#| echo: true
#| code-line-numbers: "|7"
log_harris_int <- anes_des_ind %>%
  mutate(
    EarlyVote2024 = fct_relevel(EarlyVote2024, "2. Have not voted", after = 0)
  ) %>%
  svyglm(
    design = .,
    formula = VoteHarris ~ (EarlyVote2024 + EducationGroup)^2,
    family = quasibinomial
  )
```

## Interaction effects {.smaller}

Results:

```{r}
#| echo: true
tidy(log_harris_int) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

## Interaction effects {.smaller}

```{r}
#| echo: false
tidy(log_harris_int) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

Early voting has very different associations with our outcome variable depending on the respondent's education background. For voters with less than high school education, early voters were much less likely to vote for Harris. For all other education groups, the interactions completely flip this relationship.

# Your Turn {background-image="images/header.png"}

- Go to bit.ly/MAPOR-2025-short-course
- Open `05-modeling-exercises.qmd`

```{r}
#| label: modeling-ex-time
#| echo: false
gs_ex_time <- 8
gs_ex_warn <- 3
```

```{r}
#| label: modeling-ex-clock
#| echo: false
countdown::countdown(minutes = gs_ex_time, seconds = 0, play_sound = TRUE, warn_when = gs_ex_warn * 60)
```

::: footer
bit.ly/MAPOR-2025-short-course
:::