```{r}
#| label: set-width-06
#| echo: false
options(width = 77)
```

# Sampling designs in {srvyr} {background-color='{{< brand color secondary >}}'}

## Sampling methods

Units can be selected in various ways such as:

-   **Simple random sampling (with or without replacement):** every unit has the same chance of being selected

-   **Systematic sampling:** sample individuals from an ordered list and sampling individuals at an interval with a random starting point

-   **Probability proportional to size:** probability of selection is proportional to "size"

## Complex designs

Designs can also incorporate stratification and/or clustering:

-   **Stratified sampling:** divide population into mutually exclusive subgroups (strata). Randomly sample within each stratum

-   **Clustered sampling:** divide population into mutually exclusive subgroups (clusters). Randomly sample clusters and then individuals within clusters

## Why design matters?

-   The design type impacts the variability of the estimates

-   Weights impact the point estimate and the variability estimates

-   Specifying the components of the design (strata and/or clusters) and weights in R is necessary to get correct estimates

## Determining the design {.smaller}

::: incremental
-   Check the documentation such as methodology, design, analysis guide, or technical documentation

-   Documentation will indicate the variables needed to specify the design. Look for:

    -   weight (almost always)
    -   strata and/or clusters/PSUs. Sometimes pseudo-strata and pseudo-cluster OR
    -   replicate weights (this is used instead of strata/clusters for analysis)
    -   Finite population correction or population sizes (uncommon)

-   Documentation may include SAS, SUDAAN, Stata and/or R syntax
:::

::: notes
-   We will show both ANES and RECS examples when diving into code
:::

# Designs *without* replicate weights {background-color='{{< brand color secondary >}}'}

## `as_survey_design()`: Syntax

-   Specifying the sampling design when you don't have replicate weights

-   `as_survey_design()` creates a `tbl_svy` object that then correctly calculates weighted estimates and SEs

. . .

```{r}
#| eval: false
#| code-line-numbers: "|1|2|3,7,8|4|5|6|9|"
as_survey_design(
  .data,
  ids = NULL, # cluster IDs/PSUs
  strata = NULL, # strata variables
  variables = NULL, # defaults to all in .data
  fpc = NULL, # variables defining the fpc
  nest = FALSE, # TRUE/FALSE - relabel clusters to nest within strata
  check_strata = !nest, # check that clusters are nested in strata
  weights = NULL # weight variable,
)
```

## Syntax for common designs {.smaller}

Load in the example data from {survey} package:

``` r
data(api, package="survey")
```

. . .

Simple Random Sample (SRS)

``` r
apisrs %>% as_survey_design(fpc = fpc)
```

. . .

Stratified SRS

``` r
apistrat %>% as_survey_design(strata = stype, weights = pw)
```

. . .

One-stage cluster sample with a FPC variable

``` r
apiclus1 %>% as_survey_design(ids = dnum, weights = pw, fpc = fpc)
```

. . .

Two-stage cluster sample, weights computed from population size

``` r
apiclus2 %>% as_survey_design(ids = c(dnum, snum), fpc = c(fpc1, fpc2))
```

. . .

Stratified, cluster design

``` r
apistrat %>% as_survey_design(ids = dnum, strata = stype, weights =pw, nest = TRUE)
```

::: aside
Examples from `srvyr` help documentation
:::

# Example {background-color='{{< brand color blue >}}'}

## Example: ANES 2024 {.smaller}

-   User Guide and Codebook[^1] : Section "Data Analysis, Weights, and Variance Estimation" includes information on weights and strata/cluster variables

[^1]: <https://electionstudies.org/data-center/20204time-series-study/>

> The list below indicates which weight variable to use for different types of analyses with the full release data. In general, if any post-election data are used for an estimate (alone or in combination with pre-election data), a post-election weight variable should be used, while analyses using solely pre-election data should use a pre-election weight.

## Example: ANES 2024

```{r}
#| label: design-anes-wgt-tbl
#| echo: false

table_data <- data.frame(
  Weight = c("V240101a", "V240101b", "V240102a", "V240102b", 
             "V240103a", "...", "V240108a", "V240108b"),
  description = c(
    "Pre-election fresh in-person sample alone",
    "Post-election fresh in-person sample alone",
    "Pre-election fresh web sample alone",
    "Post-election fresh web sample alone",
    "Pre-election fresh samples (fresh in-person + fresh web)",
    "...",
    "Pre-election full sample (fresh in-person + fresh web + panel)",
    "Post-election full sample (fresh in-person + fresh web + panel)"
  ),
  cluster = c("V240101c", "V240101c", "V240102c", "V240102c", 
             "V240103c", "...", "V240108c", "V240108c"),
  strata = c("V240101d", "V240101d", "V240102d", "V240102d", 
             "V240103d", "...", "V240108d", "V240108d")
)

weight_tbl <- table_data |>
  gt() |>
  cols_label(
    Weight = md("**Weight**"),
    description = md("**Use for analysis from the...**"),
    cluster = md("**variance PSU**"),
    strata = md("**variance stratum**")
  )

weight_tbl
```

## Example: ANES 2024

```{r}
#| label: design-anes-wgt-tbl2
#| echo: false

weight_tbl |>
  tab_style(cell_fill("#ff8484"),
            cells_body(rows = Weight == "V240108b"))

```

- Post-election, full sample: Weight = `V240108b`
- Use `V240108c` for variance PSU and `V240108d` for variance stratum

## Example: ANES 2024 Syntax

```{r}
#| label: design-anes-example
#| code-line-numbers: "|1|2|3|4-9|5|6|7,8"
#| eval: false

anes_des <- anes_2024 %>%
  filter(InterviewMode_Post != "3. Paper (PAPI)") %>% 
  mutate(Weight = V240108b / sum(V240108b) * 232449541) %>%
  as_survey_design(
    weights = Weight,
    strata = V240108d,
    ids = V240108c,
    nest = TRUE
  )

```

- Target population: 232,449,541
- Isabella has a blog post on calculating the population for ANES 2024 data: [https://rworks.netlify.app/posts/anes2024/](https://rworks.netlify.app/posts/anes2024/)

## Example: ANES 2024 Design

```{r}
#| label: design-anes-examp-print

summary(anes_des)
```

## National Health and Nutrition Examination Survey (NHANES) {.smaller}

-   Analysis weight: WTINT2YR
-   Variance Stratum: SDMVSTRA
-   Variance Primary Sampling Unit: VPSU
-   Large population

::: panel-tabset
### Fill in the blank

``` r
nhanes_des <- nhanes %>%
   as_survey_design(
      weights = ___________,
      ids = ___________, 
      strata = ___________, 
      fpc = ___________, 
   )
```

### Solution

``` r
nhanes_des <- nhanes %>%
   as_survey_design(
      weights = WTINT2YR,
      ids = VPSU,
      strata = SDMVSTRA,
      fpc = NULL
   )
```
:::

:::: {.fragment .tabswitch}
:::

# Designs *with* replicate weights {background-color='{{< brand color secondary >}}'}

## Replicate weights overview

- Replicate weights are a method to estimate variability that does not rely on stratum and clusters (can assist with disclosure)
- Common replicate weight methods

    -   Balanced repeated replication (BRR)
    -   Fay's BRR
    -   Jackknife
    -   Bootstrap

## `as_survey_rep()`: Syntax {.smaller}

`as_survey_rep()` creates a `tbl_svy` object that then correctly calculates weighted estimates and SEs

```{r}
#| eval: false
#| code-line-numbers: "|1|2|3|4|5|6-9|10-16"
as_survey_rep(
  .data,
  variables = NULL, # defaults to all in .data
  repweights = NULL, # Variables specifying the replication weights
  weights = NULL, # Variable specifying the analytic/main weight
  type = c(
    "BRR", "Fay", "JK1", "JKn", "bootstrap",
    "successive-difference", "ACS", "other"
  ), # Type of replication weight
  combined_weights = TRUE, # TRUE if repweights already include sampling weights, usually TRUE
  rho = NULL, # Shrinkage factor for Fay's method
  bootstrap_average = NULL, # For type="bootstrap", if the bootstrap weights have been averaged, gives the number of iterations averaged over
  scale = NULL, # Scaling constant for variance
  rscales = NULL, # Scaling constants for variance
  mse = getOption("survey.replicates.mse"), # If TRUE, compute variance based around point estimates rather than mean of replicates
  degf = NULL, # Design degrees of freedom, otherwise calculated based on number of replicate weights
)
```

## Syntax for common replicate methods {.smaller}

```{r}
#| eval: FALSE
#| code-line-numbers: "1-7|9-16|18-25|27-34"

brr_des <- dat %>%
  as_survey_rep(
    weights = WT,
    repweights = starts_with("REPWT"),
    type = "BRR",
    mse = TRUE
  )

fay_des <- dat %>%
  as_survey_rep(
    weights = WT0,
    repweights = num_range("WT", 1:20),
    type = "Fay",
    mse = TRUE,
    rho = 0.3
  )

jkn_des <- dat %>%
  as_survey_rep(
    weights = WT0,
    repweights = WT1:WT20,
    type = "JKN",
    mse = TRUE,
    rscales = rep(0.1, 20)
  )

bs_des <- dat %>%
  as_survey_rep(
    weights = pw,
    repweights = pw1:pw50,
    type = "bootstrap",
    scale = 0.02186589,
    mse = TRUE
  )
```

-   Note: this uses fake data and can't be run, just syntax examples

# Example {background-color='{{< brand color blue >}}'}

## Example: RECS 2020 {.smaller}

-   Using the microdata file to compute estimates and relative standard errors[^2]

[^2]: <https://www.eia.gov/consumption/residential/data/2020/pdf/microdata-guide.pdf>

> The following instructions are examples for calculating any RECS estimate using the final weights (NWEIGHT) and the associated RSE using the replicate weights (NWEIGHT1 – NWEIGHT60).

-   Includes R syntax for {survey} package which gets us what we need for {srvyr}

```{r}
#| eval: false
repweights <- select(RECS2020, NWEIGHT1:NWEIGHT60)
RECS <- svrepdesign(
  data = RECS2020,
  weight = ~NWEIGHT,
  repweights = repweights,
  type = "JK1",
  combined.weights = TRUE,
  scale = (ncol(repweights) - 1) / ncol(repweights),
  mse = TRUE
)
```

## Example: RECS 2020 Syntax

```{r}
#| label: design-recs-example
#| code-line-numbers: "|1|2,8|3|4|5-7"
recs_des <- recs_2020 %>%
  as_survey_rep(
    weight = NWEIGHT, 
    repweights = NWEIGHT1:NWEIGHT60, 
    type = "JK1", 
    scale = 59 / 60, 
    mse = TRUE 
  )
```

## Example: RECS 2020 Output

```{r}
#| label: design-recs-examp-print

summary(recs_des) 
```

## American Community Survey (ACS)

-   Analysis weight: PWGTP
-   replicate weights: PWGTP1-PWGTP180
-   jackknife with scale adjustment of 4/80

::: panel-tabset
### Fill in the blank

``` r
acs_des <- acs_pums %>%
   as_survey_rep(
      weights = ___________,
      repweights = ___________,
      type = ___________,
      scale = _________ 
   )
```

### Solution

``` r
acs_des <- acs_pums %>%
   as_survey_rep(
      weights = PWGTP,
      repweights = num_range("PWGTP", 1:80),
      type = "JK1",
      scale = 4/80
   )
```
:::


:::: {.fragment .tabswitch}
:::

## Tips for correctly specifying 

::: {.incremental}

- For public use datasets: **Documentation, Documentation, Documentation**

    - Read through user guides
    - Search for keywords (e.g., "weight", "strata") in the user guides
    - Google
    - Reach out to the study contact with questions
    
- For data you create: Please provide good documentation!

:::

# Thanks! {background-color='{{< brand color secondary >}}'}
