### Set-up

#### Load necessary packages

```{r}
#| label: desc-ex-set-up
#| message: false
#| warning: false
library(tidyverse)
library(srvyr)
library(srvyrexploR)
```

#### Load in data and preview it

```{r}
#| label: desc-ex-data-summary
glimpse(anes_2020)
```

Find codebooks here:

  - [ANES 2024](https://electionstudies.org/wp-content/uploads/2025/08/anes_timeseries_2024_userguidecodebook_20250808.pdf)

#### Create design objects for usage

```{r}
#| label: desc-ex-design-obj
anes_des <- anes_2024 %>%
  filter(!is.na(V240108d)) |> 
  mutate(Weight = V240108b / sum(V240108b) * 232449541) %>%
  as_survey_design(
    weights = Weight,
    strata = V240108d,
    ids = V240108c,
    nest = TRUE
  )
```

### Exercises - Part 1

1. How many females have a graduate degree (according to the ANES data)?

```{r}
# Hint: The variables `Sex` and `EducationGroup` will be useful.
```

1. What percentage of people identify as “Strong Democrat”? 

```{r}
# Hint: The variable `PartyID` indicates someone’s party affiliation.
```

1. What is the design effect for the proportion of people who voted early? 

```{r}
# Hint: The variable `EarlyVote2024` indicates whether someone voted early in 2024.
```

### Exercises - Part 2

1. What percentage of people who voted in the 2024 election identify as "Strong Republican"? 

```{r}
# Hint: The variable `VotedPres2024` indicates whether someone voted in 2020.
```

1. What percentage of people voted in both the 2024 election and the 2020 election? Include the logit confidence interval. 

```{r}
# Hint: The variable VotedPres2024 indicates whether someone voted in 2024.
```
