### Set-up

#### Load necessary packages

```{r}
#| label: set-up
#| message: false
#| warning: false
library(tidyverse)
library(survey)
library(srvyr)
library(srvyrexploR)
```

#### Load in data and preview it

```{r}
#| label: data-summary
glimpse(anes_2024)
```

Find codebook here: [ANES 2024](https://tidy-survey-r.github.io/srvyrexploR/reference/anes_2024.html)

#### Create design objects for usage

```{r}
#| label: design-obj

anes_des <- anes_2024 %>%
  filter(InterviewMode_Pre != "3. Paper (PAPI)") %>% 
  mutate(Weight = V240108b / sum(V240108b) * 232449541) %>%
  as_survey_design(
    weights = Weight,
    strata = V240108d,
    ids = V240108c,
    nest = TRUE
  )
```

### Exercises

1. Is there a relationship between party identification (`PartyID`) and whether people voted early (`EarlyVote2024`)?

```{r}
#| label: testing-exercise-1

```

2. Is there a relationship between party identification (`PartyID`) and trust in the government (`TrustGovernment`)? Use Wald as the `statistic` option.

```{r}
#| label: testing-exercise-2

```

3. Is there a difference in the distribution of gender (`Gender`) across early voting status (`EarlyVote2024`)?

```{r}
#| label: testing-exercise-3

```


